\plaintitle

### 2017-01-06 drafts/labor/piecework_lit.tex
lit rev revs interim

@@

\subsection{What is piecework? A Primer and Timeline}\label{sec:whatIsPiecework}
\topic{Aligning on--demand work with piecework requires an understanding of what piecework is.}
While it has had several definitions over time,
we can trace a constellation of characteristics that recur throughout the literature.
We'll follow this research, collecting
descriptions,
examples, and
definitions of piecework,
tracing an outline of % a working understanding of
\textit{what piecework is}.

\citeauthor{hughRaynbirdTaskWork} provides
a concise definition of piecework
(which he also calls ``measure work'', ``grate work'', and ``task work'')
by contrasting with the ``day--labourer'':
``the chief difference lies between the day--labourer,
who receives a certain some of money~\dots~for his day's work,
and the task--labourer, whose earnings depend on the quantity of work done''~\cite{hughRaynbirdTaskWork}.
\citeauthor{10.2307/2338394} defines it through examples:
``payment is made for each hectare which is pronounced to be well ploughed~\dots~for each living foal got from a mare;~\dots~for each living calf got''~\cite{10.2307/2338394}.
This framing offers a more intuitive sense;
``payment for results,'' as he calls it,
is not only common in practice, but
well--studied in labor economics as well~\cite{Figlio2007901,weitzman1976new,10.2307/3003414,BJIR:BJIR038}.

@@

\topic{We see this sense of independence in myriad times, locales, and industries.}
\citeauthor{10.2307/3827491} offers a look into the lives and culture of ``match--girls''
--- young pieceworkers, mostly women, who assembled matchsticks in the late \nth{19} century.
Of interest was their reputation ``\dots~for generosity, independence, and protectiveness,
but also for brashness, irregularity, low morality, and little education''~\cite{10.2307/3827491}.
\citeauthor{10.2307/27508091} document piecework from 1850 through 1930 in Australia,
finding similar assertions of the freedom compositors of newspapers experienced as pieceworkers:
``If a piece--work compositor~\dots~decided that he did not want to work on a particular day or night,
the management recognised his right to put a `substitute' or `grass' compositor in his place''~\cite{10.2307/27508091}.
We therefore draw a \textit{sense} of independence and autonomy inherent to piecework.
%that resonates across decades, industries, and locales where piecework is found.

@@

In summary, piecework:
\begin{inlinelist}
  \item paid workers for quantity of work done, rather than time done,
        but occasionally mixed the two payment models
  \item afforded workers freedom in when and how much to work
  \item structured tasks such that people who didn't have the training
        to engage in the traditional labor force could still participate.
\end{inlinelist}

@@

% \onlyinsubfile{\clearpage}
\subsection{A Piecework Primer}\label{sec:pieceworkPrimer} % ali adores alliteration
\msb{This section is so brief that it feels like a bit of a cheat. I think we need to frame it differently. Maybe that it should survey the different ``eras'' of piecework? And each paragraph is laying out a major historical era that we will return to later? Otherwise this entire thing feels so brief that I don't feel that I get a ton of detail out of it: it rose, people were unhappy, it fell again.}
\ali{Do you mean that we'd have (sub)subsections under this? I'm tempted to merge this subsection into the ``\namerefl{sec:whatIsPiecework}'' subsection instead. Not exactly sure how that would work (i.e. would I interleave the history or just append it, basically removing the header).}
\topic{In this section, we will discuss the history of piecework at a high level.} Our goal is to lay the groundwork necessary to address our three major questions about crowd work, rather than provide a comprehensive historical account.
% Instead, this section will attempt to provide a sense of orientation when thinking about piecework.
This section will frame piecework,
in the early days of the Industrial Revolution \&
through the political and economic turmoil of the early and mid--\nth{20} century,
and into the \nth{21} century.
% While the previous section provided a \textit{definition} of piecework,
% this section attempts to shine a light on the \textit{zeitgeist} of piecework;
% as close to a \textit{thick description} of piecework
% as we can afford~\cite{geertz1973interpretation}.
% \msb{is this really thick? as a reader it flies by really quickly and super high level...}

\topic{Piecework's history traces back further perhaps than most would expect.}
\citeauthor{grier2013computers} describes the process astronomers adopted of hiring young boys
to calculate equations in order
to better--predict the trajectories of various celestial bodies in the \nth{19} century~\cite{grier2013computers}.
George Airy was perhaps the first to rigorously apply piecework--style decomposition of tasks to work;
by breaking complex calculations into constituent parts, and
training young men to solve simple algebraic problems,
Airy could distribute work to many more people than could otherwise complete the full calculations.

\topic{Piecework may have started in the intellectual domain of astronomical calculations and projections,
but it found its foothold in manual labor.}
Piecework took hold in farm work~\cite{hughRaynbirdTaskWork},
in textiles~\cite{restructuringPieceworkBaker,riisOtherSideLives},
on railroads~\cite{Brown01041990}, and 
elsewhere in manufacturing~\cite{10.2307/3827491}.
Fordism and scientific management thrust piecework into higher gear, especially as
mass manufacturing and
a depleted wartime workforce forced industry to find new ways to eke out more production capacity.
\msb{I think a sentence here to characterize the height of piecework would be helpful --- how big a deal was it at its peak?}
\ali{I'm not sure; I'll look for some data.}
% \msb{This last sentence seems disconnected from the broader argument; cut it or draw it back to the topic sentence of the paragraph}
% \citeauthor{hart2013rise} point out that the Second World War,
% which called millions of Americans to military service,
% necessitated the rapid training and employment of
% a labor pool that hadn't historically been utilized in industrial labor: women~\cite{hart2013rise}.

% \msb{The topic sentence is about management, but the content is about discontent. Which is it?}
\topic{The early growth of piecework led to discussion surrounding how best to manage pieceworkers,
generally in an adversarial approach, consequently
frustrating workers~\cite{norton1900textile,clark1908cotton}.}
This frustrations of poor working and living conditions
(famously documented by
\citeauthor{riisOtherSideLives}~\cite{riisOtherSideLives}).
Discontent reached a crescendo when industry organizations representing
railway workers, mechanical engineers, and other industries began to speak out on pieceworkers' behalves~\cite{american1921problem,richards1904anything}.

\msb{Was the plummet related to the discontent in the prior paragraph? Placing these paragraphs next to each other seems to indicate you think so. If so, make the transition obvious.}
\topic{Piecework's popularity in the United States and Europe plummeted almost as quickly as it had climbed.}
\msb{How quickly did it plummet? By when was it gone? Where is it now?}
\msb{This sentence is empty calorie:}
\citeauthor{hart2013rise}'s work substantively explores the precipitous decline of piecework in the last third of the \nth{20} century.
\msb{Very few calorie sentence:}
In their work, \citeauthor{hart2013rise} offer a number of explanations for the sudden vanishing of piecework.
The salient suggestions include:
\begin{inlinelist}
\item the emergence of more effective, more nuanced incentive models
--- rewarding teams for complex achievements, for instance
\item the shifting of these industries (e.g., manufacturing, clothing)
to other countries
\item the quality of ``multidimensional'' work becoming too difficult to evaluate
\end{inlinelist}~\cite{hart2013rise}.

>>>>

\ali{I'm interleaving the paragraphs, so this paragraph needs to be refactored a bit to reflect that instead of suggesting that the road map will look like two distinct threads.
I'll probably try to take a shot at it in a bit, but for now don't take the wording too seriously.}

@@

\subsection{What is piecework? A Primer and Timeline}\label{sec:whatIsPiecework}

\ali{
So I'm thinking that this should flow like this:
``We'll try to understand what piecework was by tracing the development of and discourse around piecework.''
This'll give us the ability to switch back and forth between the timeline of the practice of piecework
--- that is, the ``piecework primer'' stuff down below ---
and the contemporary/concurrent research going on around the same time
(i.e. this stuff immediately following).

I'm going to interleave the paragraphs, but
if you think that's a flawed approach, then I'll bet you won't like what you'll see below that much.%
}

\topic{Aligning on--demand work with piecework requires an understanding of what piecework is.}
While it has had several definitions over the years,
we can trace a constellation of characteristics that recur throughout the literature.
We'll follow this research, collecting
descriptions,
examples, and
definitions,
tracing an outline of % a working understanding of
\textit{what piecework is}
alongside piecework's contemporary developments in practice.
\ali{This is worded a little awkwardly;
I want to say that we're looking at
the discourse about piecework at the same time that we look at
how piecework is unfolding. (wait, is that how I should word it?)}

\topic{Piecework's history traces back further perhaps than most would expect.}
\citeauthor{grier2013computers} describes the process astronomers adopted of hiring young boys
to calculate equations in order
to better--predict the trajectories of various celestial bodies in the 1830s~\cite{grier2013computers}.
George Airy was perhaps the first to rigorously apply piecework--style decomposition of tasks to work;
by breaking complex calculations into constituent parts, and
training young men to solve simple algebraic problems,
Airy could distribute work to many more people than could otherwise complete the full calculations.

\topic{Piecework may have started in the intellectual domain of astronomical calculations and projections,
but it found its foothold in manual labor.}
Piecework took hold in farm work~\cite{hughRaynbirdTaskWork},
in textiles~\cite{restructuringPieceworkBaker,riisOtherSideLives},
on railroads~\cite{Brown01041990}, and 
elsewhere in manufacturing~\cite{10.2307/3827491}.
Fordism and scientific management thrust piecework into higher gear, especially as
mass manufacturing and
a depleted wartime workforce forced industry to find new ways to eke out more production capacity.
\msb{I think a sentence here to characterize the height of piecework would be helpful --- how big a deal was it at its peak?}
\ali{I'm not sure; I'll look for some data.}

% \ali{\textbackslash topic'ing everything really makes clear when I have a run--on sentence\dots}
\topic{By \citeyear{hughRaynbirdTaskWork} we find % \citeauthor{hughRaynbirdTaskWork} offers
a concise definition of piecework
in \citeauthor{hughRaynbirdTaskWork}'s essay % 's \citetitle{hughRaynbirdTaskWork}
(where he also calls piecework ``measure work'', ``grate work'', and ``task work''),
particularly driven toward encapsulating the manual labor of farmwork.}
\citeauthor{hughRaynbirdTaskWork} does this by contrasting with the ``day--labourer'':
``the chief difference lies between the day--labourer,
who receives a certain some of money~\dots~for his day's work,
and the task--labourer, whose earnings depend on the quantity of work done''~\cite{hughRaynbirdTaskWork}.
\citeauthor{10.2307/2338394} defines it through examples:
``payment is made for each hectare which is pronounced to be well ploughed~\dots~for each living foal got from a mare;~\dots~for each living calf got''~\cite{10.2307/2338394}.
This framing offers a more intuitive sense;
``payment for results,'' as he calls it,
is not only common in practice, but
well--studied in labor economics as well~\cite{Figlio2007901,weitzman1976new,10.2307/3003414,BJIR:BJIR038}.

@@

\topic{We see this sense of independence in myriad times, locales, and industries.}
\citeauthor{10.2307/3827491} offers a look into the lives and culture of ``match--girls''
--- young pieceworkers, mostly women, who assembled matchsticks in the late \nth{19} century.
Of interest was their reputation ``\dots~for generosity, independence, and protectiveness,
but also for brashness, irregularity, low morality, and little education''~\cite{10.2307/3827491}.
\citeauthor{10.2307/27508091} document piecework from 1850 through 1930 in Australia,
finding similar notions of independence and autonomy among piecework newspaper compositors:
``If a piece--work compositor~\dots~decided that he did not want to work on a particular day or night,
the management recognised his right to put a `substitute' or `grass' compositor in his place''~\cite{10.2307/27508091}.
This sense of independence and autonomy appears to be an inherent component of piecework.
%that resonates across decades, industries, and locales where piecework is found.

\ali{I think now that this is moved to this area there's a good opportunity to frame this more as
``now that workers could choose their own schedule, style, etc\dots,
newfound interest in how to manage these workers emerged''.
Thoughts?}
\topic{The early growth of piecework led to discussion on how best to manage pieceworkers,
generally regarding workers antagonistically~\cite{norton1900textile,clark1908cotton}.}
This was a far cry from the earlier rhetoric on piecework, which promised that
piece workers would gladly work as diligently and as hard as possible because
incentive--based pay would reward workers directly for hard work~[].

This frustrations of poor working and living conditions
(famously documented by
\citeauthor{riisOtherSideLives}~\cite{riisOtherSideLives}).
Discontent reached a crescendo when industry organizations representing
railway workers, mechanical engineers, and other industries began to speak out on pieceworkers' behalves~\cite{american1921problem,richards1904anything}.

@@

% \msb{Was the plummet related to the discontent in the prior paragraph? Placing these paragraphs next to each other seems to indicate you think so. If so, make the transition obvious.}
\topic{Piecework's popularity in the United States and Europe plummeted almost as quickly as it had climbed.}
Between 1938 and 1942, piecework in a number of metal workers had climbed as steeply as from 11\% to 60\%~\cite{hart2005piecework};
by 1961, that proportion dropped to as little as 8\%~\cite{carlson1982time}.
\citeauthor{carlson1982time} details that, from 1973 to 1980, the remaining bastions piecework
--- where more than 50\% worked under incentive wage plans ---
were principally in clothes--making (e.g. hosiery, footwear, and garments).
% \msb{How quickly did it plummet? By when was it gone? Where is it now?}
% \msb{This sentence is empty calorie:}
\citeauthor{hart2013rise}'s work substantively explores the precipitous decline of piecework in the last third of the \nth{20} century.

\ali{let's try cutting it\dots
\msb{Very few calorie sentence:}
In their work, \citeauthor{hart2013rise} offer a number of explanations for the sudden vanishing of piecework.
The salient suggestions include:
\begin{inlinelist}
\item the emergence of more effective, more nuanced incentive models
--- rewarding teams for complex achievements, for instance
\item the shifting of these industries (e.g., manufacturing, clothing)
to other countries
\item the quality of ``multidimensional'' work becoming too difficult to evaluate~\cite{hart2013rise}.
\end{inlinelist}}

\ali{This feels like the Australia of paragraphs: one sentence constitutes the whole paragraph.
Should I add some analysis here?}
In summary, piecework:
\begin{inlinelist}
  \item paid workers for quantity of work done, rather than time done,
        but occasionally mixed the two payment models
  \item afforded workers freedom in when and how much to work
  \item structured tasks such that people who didn't have the training
        to engage in the traditional labor force could still participate.
\end{inlinelist}

@@

% \onlyinsubfile{\clearpage}
% \subsection{A Piecework Primer}\label{sec:pieceworkPrimer} % ali adores alliteration
% \msb{This section is so brief that it feels like a bit of a cheat. I think we need to frame it differently. Maybe that it should survey the different ``eras'' of piecework? And each paragraph is laying out a major historical era that we will return to later? Otherwise this entire thing feels so brief that I don't feel that I get a ton of detail out of it: it rose, people were unhappy, it fell again.}
% \ali{Do you mean that we'd have (sub)subsections under this? I'm tempted to merge this subsection into the ``\namerefl{sec:whatIsPiecework}'' subsection instead. Not exactly sure how that would work (i.e. would I interleave the history or just append it, basically removing the header).}
\ali{there's no need for this paragraph:

\topic{In this section, we will discuss the history of piecework at a high level.} Our goal is to lay the groundwork necessary to address our three major questions about crowd work, rather than provide a comprehensive historical account.
% Instead, this section will attempt to provide a sense of orientation when thinking about piecework.
This section will frame piecework,
in the early days of the Industrial Revolution \&
through the political and economic turmoil of the early and mid--\nth{20} century,
and into the \nth{21} century.}
% While the previous section provided a \textit{definition} of piecework,
% this section attempts to shine a light on the \textit{zeitgeist} of piecework;
% as close to a \textit{thick description} of piecework
% as we can afford~\cite{geertz1973interpretation}.
% \msb{is this really thick? as a reader it flies by really quickly and super high level...}

% \msb{This last sentence seems disconnected from the broader argument; cut it or draw it back to the topic sentence of the paragraph}
% \citeauthor{hart2013rise} point out that the Second World War,
% which called millions of Americans to military service,
% necessitated the rapid training and employment of
% a labor pool that hadn't historically been utilized in industrial labor: women~\cite{hart2013rise}.

% \msb{The topic sentence is about management, but the content is about discontent. Which is it?}
